Schoenfeld residuals

### 2022-12-09 text/22/definition-schoenfeld-residuals.Rmd

We want to find "good values" of $\lambda$, ones that are associated with large (and therefore more likely) values of the density. The maximum likelihood estimate is the value of $\lambda$ that maximizes the products of all these densities. For the exponential case, this would be

$L(\lambda, Y, X)=\prod_i (\lambda_0+\lambda_1X_i) exp(-Y_i/(\lambda_0+\lambda_1X_i))$

Let's make up some simple data for this example. Let Y=(1, 5, 2) and X=(0, 0, 1). The likelihood function in this simple case is

$L(\lambda, Y, X)=\lambda_0 exp(-1 \lambda_0) \lambda_0 exp(-5 \lambda_0) (\lambda_0+\lambda_1) exp(-5(\lambda_0+\lambda_1))$

@@

```{r}
n0 <- 20
n1 <- 21

u0 <- seq(1/n0, 0.7, length=n0)
u1 <- seq(0, 1.4, length=n1)

l <- function(u0, u1) {
  2*log(u0)-9*u0+log(u0+u1)-2*u1
}

l_matrix <- matrix(nrow=n0, ncol=n1)

for (i0 in 1:n0) {
  for (i1 in 1:n1) {
    l_matrix[i0, i1] <- l(u0[i0], u1[i1])
  }
}

plot(u0, l_matrix[1:n0, 1])
plot(u1, l_matrix[1, 1:n1])
persp(u0, u1, l_matrix, theta=15, phi=30)

cycle_fast <- function(g=1001, cmin=0, cmax=1, rep=g) {
  cyc <- seq(cmin, cmax, length=g)
  return(rep(cyc, times=rep))
}

cycle_slow <- function(g=1001, cmin=0, cmax=1, rep=g) {
  cyc <- seq(cmin, cmax, length=g)
  return(rep(cyc, each=rep))
}
cycle_fast(3)
cycle_slow(3)

#  persp3D(z=d, xlab="lambda1", ylab="lambda2",zlab="Density", zlim=c(0, zmax),
#          theta=th, phi=ph, colvar=NULL, col="white", border="black",
#          bty="u", col.axis="gray90")

```

@@

The maximum value for the log likelihood is `r max(l_matrix)`. If you set $\lambda_2=0$, the log_likelihood is `r max(l_matrix[ , 1])`. Since these values are close, you can consider the hypothesis to be reasonable.

@@

>>>>

We want to find "good values" of $\lambda$, ones that are associated with large (and therefore more likely) values of the density. The maximum likelihood estimate is the value of $\lambda$ that maximizes the products of all these densities.

Let's make up some simple data for this example. Let Y=(3, 5, 1) and X=(0, 0, 1). The likelihood function in this simple case is

$L(\lambda, Y, X)=\lambda_0 exp(-3 \lambda_0) \lambda_0 exp(-5 \lambda_0) (\lambda_0+\lambda_1) exp(-(\lambda_0+\lambda_1))$

@@

```{r}
library(plot3D)

l <- function(lambda0, lambda1) {
  2*log(lambda0)-9*lambda0+log(lambda0+lambda1)-lambda1
}

n0 <- 50
n1 <- 50

lambda0 <- seq(1/n0, 1, length=n0)
lambda1 <- seq(0, 1, length=n1)

fine_mesh <- matrix(nrow=n0, ncol=n1)

for (i0 in 1:n0) {
  for (i1 in 1:n1) {
    fine_mesh[i0, i1] <- l(lambda0[i0], lambda1[i1])
  }
}

persp3D(
  x=lambda0, 
  y=lambda1, 
  z=fine_mesh, 
  zlim=c(min(fine_mesh), -5),
  contour=TRUE,
  theta=15,
  phi=30, 
  xlab="lambda0",
  ylab="lambda1",
  zlab="log likelihood")
segments3D(
  x0=0.25,
  y0=0.75, 
  z0=min(fine_mesh),
  x1=0.25,
  y1=0.75,
  z1=max(fine_mesh),
  add=TRUE)
text3D(
  x=0.25,
  y=0.75, 
  z=min(fine_mesh),
  labels=round(max(fine_mesh), 1),
  col="red",
  add=TRUE)
segments3D(
  x0=0.3333,
  y0=0, 
  z0=min(fine_mesh),
  x1=0.3333,
  y1=0,
  z1=max(fine_mesh[ , 1]),
  add=TRUE)
text3D(
  x=0.3333,
  y=0., 
  z=min(fine_mesh),
  labels=round(max(fine_mesh[ , 1]), 1),
  col="red",
  add=TRUE)
```

@@

The maximum value for the log likelihood is `r max(fine_mesh)`. If you set $\lambda_2=0$, the log_likelihood is `r max(fine_mesh[ , 1])`. Since these values are close, you can consider the hypothesis to be reasonable.

@@

To find out where the maximum occurs, take the first two partial derivatives and set them equal to zero.

$\frac{\partial /l}{\partial \lambda0} = 2/\lambda_0 -9 + 1/(\lambda_0+\lambda_1) = 0$

$\frac{\partial /l}{\partial \lambda1} = 1/(\lambda_0+\lambda_1) - 1 = 0$

The second equation tells you that

$\lambda_0+\lambda_1=1$

Substituting that into the second equation gives you

$\lambda_0=0.25$

and then you can show that

$\lambda_1=0.75$

>>>>>>> 17772f400cc3acd5b2c61623fc3cf2c98df50124
